---
title: "The Truth is Out There"
author: "Katie Press"
description: "Creating an X-Files episode dataset using web scraping."
categories: ["R", "syntax"]
tags: ["rmarkdown", "css"]
date: "2021-01-01"
images:
- alt: X Files Chart 1
  src: img/2021/thetruth1_plot1.png
  stretch: horizontal
output: 
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(message = FALSE, warning = FALSE, show=TRUE)
library(tidyverse)
library(rvest)
library(janitor)
library(lubridate)
library(ggiraph)
library(extrafont)
library(tidytext)
```


I'm a huge X-Files fan. I also love working with text data. I've had an idea floating around for a while to do a tidy text analysis on X-Files episode scripts. The X-Files has been around long enough that there are tons of fandom sites, and you can easily find transcripts of the original 9 seasons.First, I wanted to get some basic information about the episodes, so that's what this post will focus on. My first thought was to go to Wikipedia. There is a page with tables for each season and I can use that as the base URL for scraping. 

Packages used in this first episode:

- Tidyverse, obviously. This is always the first package I load. 
- Janitor, which has a couple of functions I like to use, especially clean_names() to clean and remove special characters from column names in new datasets. 
- Rvest, which can be used to scrape data from websites. 
- Googlesheets4, which is an update of the original googlesheets package. I can use this to store my data because I have more than one computer I use on a regular basis.  
- Extrafont (pretty self-explanatory).
- Ggiraph for graph animation.

Now, on to the X-Files.


```{r}
data(mtcars)

head(mtcars)
```

